---
title: "ESM206 Lab 8 KS"
author: "Katelin Seeto"
date: "11/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

install.packages("ggbeeswarm")
install.packages("car")

library(tidyverse)
library(car) # Just use for the Type III ANOVA check here
library(janitor)
library(ggbeeswarm)
library(kableExtra)
```

### Read in data (penguins.csv)
```{r}
penguins <- read_csv("penguins.csv") %>% 
  janitor::clean_names()
```
###2. Find counts of penguins by sex and species
can use count to know how many penguins are in each group

```{r}
penguin_counts <- penguins %>% 
  count(species, sex)

penguin_counts
```
### 3. Compare flipper size for female penguins across the 3 spp
For single factor ANOVA:

What is single factor?
Species

Flipper length= single variable

```{r}
penguin_f <- penguins %>% 
  filter(sex=="FEMALE")

# Faceted histogram
ggplot(data= penguin_f, aes(x= flipper_length_mm)) +
  geom_histogram(aes(fill = species),
                 alpha = 0.5,
                 show.legend = FALSE,
                 bins = 15) +
  facet_wrap(~species, scales = "free")

#Faceted QQ Plot
ggplot(data= penguin_f, aes(sample = flipper_length_mm)) +
  geom_qq(aes(fill = species),
                 alpha = 0.5,
                 show.legend = FALSE,
                 bins = 15) +
  facet_wrap(~species, scales = "free")


```

###4. Find summary statistics
Does satisfy assumptions of ANOVA? 
Assume that variances of groups comparing are the same
Will do hypothesis test to test for equal variances

```{r}
#Creating summary tables

penguin_f_summary <- penguin_f %>% 
  group_by(species) %>% 
  summarize(mean_flipper_length = mean(flipper_length_mm),
            sd_flipper_lenth= sd(flipper_length_mm),
            sample_size = n(),
            se_flipper_length =sd(flipper_length_mm)/ sqrt(n()),
            var_flipper_length = var(flipper_length_mm)
  )
            
```

Levene's Test: (hypothesis test for equal variances)

```{r}
leveneTest(flipper_length_mm ~ species, data= penguin_f)
```

null hypothesis for Levene's test is that group variances are equal
Based on this p val of 0.057 would I accept or reject of significance level is 0.05?
Would retain null hyp that variances are equal
  is close enough
  
Conclusion: I've done exploratory analysis and feel that a one-way ANOVA is an appropriate way to compare means between the 3 spp. 

###5. Make a beeswarm plot with mean and sd overlayed on top of it

The 3 Most common error bars:
- standard deviation
- standard error (most commonly shown, but figure caption should explain)
- confidence interval


```{r}
ggplot() +
  geom_beeswarm(data= penguin_f,
                aes(x= species, y= flipper_length_mm),
                size= 1,
                alpha= 0.6,
                color= "gray50")+
  scale_x_discrete(labels= c("Adelie", "Chinstrap", "Gentou"))+
  geom_point(data= penguin_f_summary, aes(x= species, y= mean_flipper_length), 
             color= "red",
             size= 2)+
  geom_errorbar(data= penguin_f_summary,
                aes(x= species,
                    ymin= mean_flipper_length- sd_flipper_lenth, 
                    ymax = mean_flipper_length + sd_flipper_lenth),
                color = "red",
                width = 0.1) +
  theme_light()

```
**Note** Make sure if you're showing summary statistics, you actually say what you're showing
**Figure 1.** Flipper length (mm) for three female penguin species (Adelie, Chinstrap, Gentou) observations at Palmer station LTER, Antarctica. Gray points indicate individual fipper lengths, red dots are group means, error bars indicate 1 standard deviation. Data Source


###6. One-Way ANOVA
```{r}
penguin_aov <- aov(flipper_length_mm ~ species, data= penguin_f)
summary(penguin_aov)

#Note Large F value= variance across groups is very large compared to in group spread
#Do you think is likely samples are drawn from groups with same mean? No, unlikely these samples are drawn from pop with same mean
```
H0 for omnibus ANOVA: Means are equal across all groups
HA: Means are NOT equal across all groups (can mean at least 2 differ, or all differ)

Conclusion: Mean flipper lengths for the 3 spp of female penguins are not equal

